---
title: "Introduction to Synder"
author: "Zebulun Arendsee"
output:
    rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Introduction to Synder}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

Other vignettes:

 - introduction to Synder
 - [algorithmic details](algorithm.html)
 - [a case study in yeast](yeast.html)

The primary function of synder is to map genomic intervals of interest from a
query genome to search intervals in the target genome. In this vignette, I
will showcase the basic functionallity using a toy data set.

## Installation

Synder is available from github, installation requires devtools:

```
devtools::install_github('arendsee/synder')
```

```{r, echo=TRUE}
# imports required for this vignette
library(synder)
library(knitr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(readr)
```

## Mapping intervals on the simple linear synteny map 

Sample raw data, as `data.frames` rather than the dedicated Bioconductor types,
can be found in `intro_1`, `intro_2` and `toy`.

```{r}
data(intro_1)
```

This dataset contains the following synteny map, which can be converted to a
data.frame with `as.data.frame`.

```{r, echo=TRUE, results='asis'}
knitr::kable(as.data.frame(intro_1$synmap))
```

```{r, echo=TRUE}
(syn <- intro_1$synmap)
```

The synmap object has specialized print and plot functions.

```{r, echo=TRUE}
class(syn)
```

The print function is curtesy of the inherited `GRangePairs` class

```{r, echo=TRUE}
print(syn)
```

The plot function shows the synteny map as an alignment.

```{r, echo=TRUE, fig.show='hold', fig.cap="Synteny map plot"}
plot(syn)
```

We can map intervals from query to target interactively with the `anon_search`
function.

```{r, echo=TRUE, fig.show='hold', fig.cap="Finding query in a syntenic hole"}
x <- anon_search(syn, 240, 260, 'que')
plot(x, syn)
```

The plot function here takes the search results and plots them in the context
of the synteny map. The x-axis shows position relative to the query, and the
y-axis relative to the target. The `anon_search` and `search` functions map
between th axes.

Notice that the search interval snaps to the size of the gap in the target
synteny map.

```{r, echo=TRUE, fig.show='hold', fig.cap="Finding query where one side overlaps"}
x <- anon_search(syn, 240, 310, 'que')
plot(x, syn)
```

Both sides snap to the nearest target edge.

## Inversions and contiguous sets

`synder` breaks synteny maps into contiguous sets of syntenic intervals. The
sets are defined as sequences of syntenic intervals that are adjacent on both
the query and target side and are not overlapping.

```{r}
data(intro_2)
plot(intro_2$synmap)
plot(synder::dump(intro_2$synmap))
```
