---
title: "Yeast Ty1 mobile element case study"
author: "Zebulun Arendsee"
date: "`r Sys.Date()`"
output:
    rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Ty1 case study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

Synder traces orthology independent of sequence similarity. This allows unique
mappings of sequences even when they are members of large families.

The `ty1` dataset includes synteny maps built between *S. cerivisiae* and 8
other yeast species. These maps were built using mummer.

```{r}
data(ty1)
```

```{r}
names(ty1$glens)
```

```{r}
lapply(ty1$glens, nrow)
```


# Map density

```{r}
syn <- list()
syn$c500 <- lapply(ty1$synmaps, subset, (qstop - qstart + 1) > 500)
syn$c400 <- lapply(ty1$synmaps, subset, (qstop - qstart + 1) > 400)
syn$c300 <- lapply(ty1$synmaps, subset, (qstop - qstart + 1) > 300)
syn$c200 <- lapply(ty1$synmaps, subset, (qstop - qstart + 1) > 200)
syn$c100 <- lapply(ty1$synmaps, subset, (qstop - qstart + 1) > 100)
syn$c0   <- lapply(ty1$synmaps, subset, (qstop - qstart + 1) > 0)
```

```{r, eval=FALSE}
# Preparing glens
glens_f = list.files(pattern='.*.glen')
glens_names <- sub('\\..*', '', glens_f)
glens <- lapply(glens_f, synder::read_conlen)
names(glens) <- glens_names
```

```{r}
results <- list()
results$c500 <- lapply(syn$c500, function(x) search(x, ty1$gff))
results$c400 <- lapply(syn$c400, function(x) search(x, ty1$gff))
results$c300 <- lapply(syn$c300, function(x) search(x, ty1$gff))
results$c200 <- lapply(syn$c200, function(x) search(x, ty1$gff))
results$c100 <- lapply(syn$c100, function(x) search(x, ty1$gff))
results$c0   <- lapply(syn$c0  , function(x) search(x, ty1$gff))
```

Now we want to infer orthology independent of sequence similarity. So I will
subtract the query intervals from the synteny maps.


```{r}
maps <- lapply(ty1$synmaps, function(x) map(x, ty1$gff))

holy <- lapply(
    ty1$synmaps,
    function(x){
        m <- map(x, ty1$gff)
        m_str <- paste0(m$qcon, m$qstart, m$qstop)  
        s_str <- paste0(x$qcon, x$qstart, x$qstop)
        idx <- s_str %in% m_str
        x[!idx, ]
    }
)

holy_results <- lapply(holy, function(x) search(x, ty1$gff))

holy_results_k100 <- lapply(holy, function(x) search(x, ty1$gff, k=100))
```
